A Descriptive Analysis of How Psychology Journals Handle Post-Publication Critiques

### 2023-12-26 manuscript_NHB.Rmd
remove .png files update manuscript
Removed unnecessary images. Updated manuscript to version that knits successfully

---
title: A Descriptive Analysis of How Psychology Journals Handle Post-Publication
  Critiques
format:
  pdf: default
  journal: "sn-nature-human-behaviour"
  keep-tex: true
author:
  - name: Annie Whamond
    affiliations:
      - name: University of Melbounre
        id: 1
        department: Melbourne School of Psychological Sciences
        address: 161 Barry St
        city: Carlton
        state: VIC
        postal-code: 3010
    email: whamonda@unimelb.edu.au
    corresponding: true
  - name: Simine Vazire
    affiliations:
      - ref: 1
  - name: Beth Clarke
    affiliations:
      - ref: 1
  - name: Nicholas Moodie
    affiliations:
      - ref: 1
  - name: Sarah Schiavone
    affiliations:
      - ref: 1
  - name: Robert T. Thibault
    affiliations:
      - name: Stanford University
        id: 2
        department: Meta Research Innovation Centre
        address: 450 Jane Stanford Way
        city: Stanford
        state: CA
        postal-code: 94305
  - name: Tom E. Hardwicke
    affiliations:
      - ref: 1
abstract: >
  Fostering critical discussion is essential to upholding the ethos of
  science as an unbiased, collaborative, and self-correcting endeavor. Journals
  can support this by offering options for post-publication critique, such as
  commentaries or letters, allowing peers to voice substantiated concerns of
  empirical articles. We provide a descriptive, cross-sectional analysis of
  post-publication critique policies and prevalence across both prominent
  randomly-selected and psychology journals. We found that, overall, the
  majority of psychology journals do not offer explicit options for
  post-publication critique. When they did, length and time-to-submit limits
  were often not stated. Critique prevalence was rare with only a single event
  identified. Given the widespread recognition of the replication crisis in
  psychology, allowing avenues for criticisms to be voiced may help increase the
  credibility within the field. To support transparency and self-correction,
  psychology journals should offer clearly stated post-publication critique
  options and encourage their use.
keywords:
  - metascience
  - post-publication cretique
  - psychology
  - credibility
bibliography: bibliography.bib

---

```{r}
library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
```

```{r}
source(here('analysis','functions.R')) # load custom functions
source(here('analysis','raw_to_primary.R')) # converts raw data to primary data
```

```{r}
# Load processed data files

d_articles <- read_csv(here('data','processed','d_articles.csv'), show_col_types = F) # article data 
d_journals <- read_csv(here('data','processed','d_journals.csv'), show_col_types = F) # journal data 
```

```{r}
# Create separate tibbles for Prominent and Randomly-selected journal samples

dj_prominent <- d_journals %>%
  filter(sample_id != "random") %>%
  filter(is_empirical == TRUE)

dj_random <- d_journals %>%
  filter(sample_id != "prominent") %>%
  filter(is_empirical == TRUE)
```

# Main

@@

```{r journal-characteristics}

# below we ascertain journal JIF characteristics

promJournalCharacteristics <- dj_prominent %>% 
  filter(ppc_type == "A") %>%
  summarise(promMedian_jif = round(median(jif),2),
            promMin_jif = round(min(jif),2),
            promMax_jif = round(max(jif),2),
            promIQR_jif = round(IQR(jif),2)
  )

@@

```{r, echo = FALSE}
# Filter and transform restrictions for ppc_length
prominent_length_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>% # removes Archive Only PPC types
  mutate(ppc_length = case_when(
    ppc_length %in% c(1000, 1100, 1200, 1500, 2000, 2500, 2650, 3000, 3500, 400, 500, 5000, 750, 800, 8000) ~ 'Quantitative',
    ppc_length == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all length limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>% 
  group_by(ppc_length) %>%
  summarise(count = n())

# Filter and transform restrictions for ppc_time
prominent_time_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>%
  mutate(ppc_time = case_when(
    ppc_time %in% c(13.05, 26.1, 39.15, 4, 52) ~ 'Quantitative',
    ppc_time == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all time limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>%
  group_by(ppc_time) %>%
  summarise(count = n())

@@

# Create the second graph
graph2 <- dj_prominent %>%
  filter(ppc_type == "A") %>% # Keeps only first PPC row for each journal
  filter(!is.na(has_ppc)) %>%
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Archive',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

@@

```{r}
# Length restriction statement types of explicit PPC types (prominent sample)

@@

# Peer-review statements of explicit PPC types (prominent sample)
prominent_review_data <- dj_prominent %>%
  filter(!is.na(ppc_review)) %>% # keep only entries that identify PPC option
  filter(has_ppc != "ARCHIVE") %>% # remove implicit yes examples
  mutate(ppc_review = case_when(
    ppc_review == 'YES' ~ 'Yes',
    ppc_review == 'NO' ~ 'No',
    ppc_review == "NOT STATED" ~ "Not Stated"
  )) %>%
  group_by(ppc_review) %>%
  summarise(count = n()) %>%
  ungroup()

@@

## Randomly-selected Psychology Journals
```{r journal-characteristics}

# below we ascertain journal JIF characteristics

@@

```
Amongst our sample of 100 randomly-selected psychology journals, Journal Impact Factor data was unavailable for 25 journals. Median 2021 Journal Impact Factor was `r randJournalCharacteristics %>% pull(randMedian_jif)` (IQR = `r randJournalCharacteristics %>% pull(randIQR_jif)`, range = `r randJournalCharacteristics %>% pull(randMin_jif)` - `r randJournalCharacteristics %>% pull(randMax_jif)`). Figure *XX* shows the observed journal policies and, of the explicitly offered critique options identified, how restrictions on length and time-to-submit were stated.

### Figure *XX*: *Post-Publication Critique Policy and Restriction Statements in Randomly-selected Psychology Journals*
```{r}
# Create and combine restiction statement graphs with ppc acceptance (random sample)

@@

```
*Note.* PPC = post-publication critique. Orange bar (left) indicates policy statements for each journal included in sample (n = 100). Blue bars (right) indicate how length and time-to-submit restrictions were stated in cases where journals had explicit post-publication critique policies in place (n = 25).

```{r}
# Length restriction statement types of explicit PPC types (random sample)

@@

# Discussion

# Data Availability

# Code Availability

>>>>

---
title: "A Descriptive Analysis of How Psychology Journals Handle Post-Publication Critiques"
author: "Whamond, A., Vazire, S., Clarke, B., Moodie, N., Schiavone, S., Thibault, R. T., & Hardwicke, T. E."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include = FALSE}
renv::activate()

library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(ggrepel) # for ggplot text labels
library(patchwork) # for multipanel plot layouts
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
```

```{r echo=FALSE, message=FALSE}
# Load processed data files

d_articles <- read_csv(here('ProjCode', 'GitClone', 'PROJ_2023_psych-journal-critique', 'data','processed','d_articles.csv'), show_col_types = F) # article data 
d_journals <- read_csv(here('ProjCode', 'GitClone', 'PROJ_2023_psych-journal-critique', 'data','processed','d_journals.csv'), show_col_types = F) # journal data 
```

```{r echo=FALSE, message=FALSE}
# Create separate tibbles for Prominent and Randomly-selected journal samples

dj_prominent <- d_journals %>%
  filter(sample_id != "random") %>% # remove entries from the random sample
  filter(is_empirical == TRUE) # remove ineligible (non-empirical) journals

dj_random <- d_journals %>%
  filter(sample_id != "prominent") %>% # remove entries from the prominent sample
  filter(is_empirical == TRUE) # remove ineligible (non-empirical) journals
```

# Main
*Add 'Introduction' section*

@@

```{r prominent journal-characteristics, echo=FALSE, message=FALSE}

# Ascertain prominent journal JIF characteristics

promJournalCharacteristics <- dj_prominent %>% 
  filter(ppc_type == "A") %>%       # Keep only single entry for each journal
  summarise(promMedian_jif = round(median(jif),2), 
            promMin_jif = round(min(jif),2),
            promMax_jif = round(max(jif),2),
            promIQR_jif = round(IQR(jif),2)
  )

@@

```{r, echo = FALSE, message=FALSE}
# Filter and transform restrictions for ppc_length
prominent_length_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>% # removes 'Archive Only' PPC types
  mutate(ppc_length = case_when(
    ppc_length %in% c(1000, 1100, 1200, 1500, 2000, 2500, 2650, 3000, 3500, 400, 500, 5000, 750, 800, 8000) ~ 'Quantitative',
    ppc_length == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all length limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>% 
  group_by(ppc_length) %>%
  summarise(count = n())

# Filter and transform restrictions for ppc_time
prominent_time_data <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  filter(has_ppc != "ARCHIVE") %>% # removes 'Archive Only' PPC types
  mutate(ppc_time = case_when(
    ppc_time %in% c(13.05, 26.1, 39.15, 4, 52) ~ 'Quantitative',
    ppc_time == "NOT STATED" ~ "Not Stated",
    TRUE ~ 'Qualitative' # Summarises all time limits as 'Quantitative', 'Qualitative', or 'Not Stated'
  )) %>%
  group_by(ppc_time) %>%
  summarise(count = n())

@@

# Create the second graph
graph2 <- dj_prominent %>%
  filter(ppc_type == "A") %>% # Keeps only first entry for each journal
  filter(!is.na(has_ppc)) %>% # Remove journal rows with no (NA) PPC options 
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Archive',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

@@

```{r echo=FALSE, message=FALSE}
# Length restriction statement types of explicit PPC types (prominent sample)

@@

# Peer-review statements of explicit PPC types (prominent sample)
prominent_review_data <- dj_prominent %>%
  filter(!is.na(ppc_review)) %>% # keep only entries that identify PPC option
  filter(has_ppc != "ARCHIVE") %>% # remove implicit yes examples
  mutate(ppc_review = case_when(
    ppc_review == 'YES' ~ 'Yes',
    ppc_review == 'NO' ~ 'No',
    ppc_review == "NOT STATED" ~ "Not Stated")
  ) %>%
  group_by(ppc_review) %>%
  summarise(count = n()) %>%
  ungroup()

@@

## Randomly-selected Psychology Journals
```{r randomly-selected journal-characteristics, echo=FALSE, message=FALSE}

# Ascertain randonly-selected journal JIF characteristics

@@

```
Among our sample of 100 randomly-selected psychology journals, Journal Impact Factor data was unavailable for 25 journals. Median 2021 Journal Impact Factor was `r randJournalCharacteristics %>% pull(randMedian_jif)` (IQR = `r randJournalCharacteristics %>% pull(randIQR_jif)`, range = `r randJournalCharacteristics %>% pull(randMin_jif)` - `r randJournalCharacteristics %>% pull(randMax_jif)`). Figure *XX* shows the observed journal policies and, of the explicitly offered critique options identified, how restrictions on length and time-to-submit were stated.

### Figure *XX*: *Post-Publication Critique Policy and Restriction Statements in Randomly-selected Psychology Journals*
```{r echo=FALSE, message=FALSE}
# Create and combine restriction statement graphs with PPC acceptance (random sample)

@@

*Note.* PPC = post-publication critique. Orange bar (left) indicates policy statements for each journal included in sample (n = 100). Blue bars (right) indicate how length and time-to-submit restrictions were stated in cases where journals had explicit post-publication critique policies in place (n = 25).

```{r echo=FALSE, message=FALSE}
# Length restriction statement types of explicit PPC types (random sample)

@@

# Discussion
*Add 'Discussion' section*

# Methods
*Add 'Methods' section*

# Data Availability
The data used for this paper are available from the project’s OSF repository (link: [https://osf.io/8k7m4/](https://osf.io/8k7m4/))

# Code Availability
The analysis code for all results reported in this paper are available from the project’s OSF repository (link: [https://osf.io/8k7m4/](https://osf.io/8k7m4/))
